---
title: "Covid91 vaccine study Final2022"
author: "Diego Martinez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
Males = filter(FinalData, sex == "M")
Females = filter(FinalData, sex == "F")
LGBTQ  = filter(FinalData, LGBTQ == "gay")
DrugUser = filter(FinalData, DrugUser == "yes")
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## General Null Hypothesis & Alternative Hypothesis, & Methods

$H_0$: There is no difference between how effective the vaccine is for whichever sub-group uses it. These two values are independent

$H_A$: There is a difference between how effective the vaccine is for whichever sub-group uses it. These two values are dependent. 

This statistical final is CAT~CAT since both values are categorical. Whether the vaccine is effective (or not) can be categorized and so can the sub-groups. Thus, I will be using the following strategies: A bar chart (graphical descriptive results), a data table (numerical descriptive results), a chi-squared test (inferential results), and a fisher exact test (inferential descriptive results) for all 4 of the sub-groups. After I've done this for all 4 sub-groups, I will separately consider the vaccine for each sub-group & give my recommendation (based off of the statistics) whether or not each sub-group should take the vaccine. This final will be broken down into 4 questions, 4 null & alternative hypothesis, 4 brief conclusions, & 1 final (very descriptive) conclusion. This general null hypothesis & alternative hypothesis will not be answered. This is to better organize my study. My next 4 null & alternative hypothesis's will be answered as required. 

## Males

### Null & Alterantive Hypothesis, & Methods (Question #1)

$H_0$: There is no correlation between males & the vaccine. The vaccine doesn't work on males: you have the same likelihood of contracting covid-19 if you do or do not take the vaccine (if you are a male).

$H_A$: There is a correlation between males & the vaccine. The drug is either beneficial or disadvantageous in preventing males from contracting covid-19. You don't have the same likelihood (if you do or do not take the vaccine) of contracting the disease.

I'll use a bar chart for graphical descriptive results. I'll use a numerical data table for numerical descriptive results. I'll use a chi-squared test & a fisher exact test for inferential results. Than lastly, conclude if I fail to reject or reject the null hypothesis. 

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment, data = Males)
barchartGC(~infected + treatment, data = Males, type = "percent")
```

The vaccine appears to make males less susceptible to covid-19. Since in the first bar chart (in frequency), you can visibly tell that males who do take the vaccine contract covid-19 less often. My second bar chart (in percentage), only furthers proves my claim. Since the bar chart value (for infected) decreases in the covid-19 department when males do take the vaccine by about 40% (if the placebo infected rate is at 70% & if the vaccine infected rate is at 30%). Males so far, have a positive result when they take the vaccine. 

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment, data = Males)
rowPerc(table1)
colPerc(table1)
```

My (row percentages) data table states that my initial estimate for the decreased likelihood of males to contract covid-19 with the vaccine being 40% is valid. Since the data table shows that only 32% of the infected population is vaccinated. Thus, when I subtract the two percentages, the placebo percentage is higher by 35.56%. Quite close to my estimate of 40%.  My (column percentages) data table shows that only 1.28% of males (who are vaccinated) have contracted covid-19. Thus, 1.35% less of the vaccinated male population is infected with covid-19 compared to those who aren't vaccinated. Since when I subtract the two percentages, the placebo percentage is higher by 1.35%. Males, so far, continuously show to have a positive result when they take the vaccine. 

### Inferential Descriptive Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The null hypothesis has to be rejected since the chi-squared test states that the p-value is 0.00000000004812. This number is way below than the .05 mark. The chi-squared testGC also rejects the null hypothesis. It has a chi-squared statistic of 43.25 with a degrees of freedom of 1.  We reject the null hypothesis since it states that the p-value is 0. Also way below .05. The fisher exact test states that the p-value is 0.00000000002645. The fisher exact test also states that the odds ratio is 2.09 to 1. Our alternative hypothesis states that the odds ratio is not equal to 1. Since all 3 p-values are below .05 & the odds ratio contradicts the null hypothesis, I can safely reject the null hypothesis. There is a difference, it being a beneficial one for males. 

### Males Discussion & Conclusion

The original question is if the vaccine is either beneficial, disadvantageous, or neutral in preventing males from contracting covid-19. I can conclude that there IS a difference. I can REJECT the null hypothesis. Since all 3 p-values were below .05 & since the odds ratio contradicts the null hypothesis. Now than, the vaccine POSITIVELY effects males. When males take the vaccine, their odds of contracting covid-19 DECREASES. Thus, I DO recommend that males do take the vaccine since it helps decrease their risks of contracting the deadly disease. 


## Females

### Null Hypothesis & Alternative Hypothesis, & Methods (Question #2)

$H_0$: There is no correlation between females & the vaccine. The vaccine doesn't work on females: you have the same likelihood of contracting covid-19 if you do or do not take the vaccine (if you are a female).

$H_A$: There is a correlation between females & the vaccine. The drug is either beneficial or disadvantageous in preventing females from contracting covid-19. You don't have the same likelihood (if you do or do not take the vaccine) of contracting the disease.

I'll use a bar chart for graphical descriptive results. I'll use a numerical data table for numerical descriptive results. I'll use a chi-squared test & a fisher exact test for inferential results. Than lastly, conclude if I fail to reject or reject the null hypothesis. 

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment, data = Females)
barchartGC(~infected + treatment, data = Females, type = "percent")
```

The vaccine appears to make females less susceptible to covid-19. Since in the first bar chart (in frequency), you can visibly tell that females who do take the vaccine contract covid-19 less often. My second bar chart (in percentage), only furthers proves my claim. Since the bar chart value (for infected) decreases in the covid-19 department when females do take the vaccine by about 20% (if the placebo infected rate is at 60% & if the vaccine infected rate is at 40%). Females so far, have a positive result when they take the vaccine. 


### Numerical Descriptive Results

```{r}
table2 <- xtabs(~infected + treatment, data = Females)
rowPerc(table2)
colPerc(table2)
```

My (row percentages) data table shows that my initial estiamte for the decreased likelihood of females to contract covid-19 with the vaccine being 20% is valid. Since the data table shows that only 41% of the infected population is vaccinated. Thus, when I subtract the two percentages, the placebo percentage is higher by 17.78%. Quite close to my estimate of 20%. My (column percentages) data table shows that only 1.62% of females (who are vaccinated) have contracted covid-19. Thus, .68% less of the vaccinated female population is infected with covid-19 compared to those who aren't vaccinated. Since when I subtract the two percentages, the placebo percentage is higher by .68%. Females, so far, continuously show to have a positive result when they take the vaccine. 


### Inferetial Descriptive Results 

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The null hypothesis has to be rejected since the chi-squared test states that the p-value is .0011. This number is way lower than the .05 mark. The chi-squared testGC also rejects the null hypothesis. It has a chi-squared of 10.65 with a degrees of freedom of 1. We reject the null hypothesis since it states that the p-value is .0011. Also way below .05. The fisher exact test states that the p-value is 0.0009484. The fisher exact test also states that the odds ratio is 1.43 to 1. Our alterantive hypothesis states that the odds ratio is not equal to 1. Since all 3 p-value are below .05 & the odds ratio contradicts the null hypothesis, I can safely reject the null hypothesis. There is a difference, it being a beneficial one for females. 

### Females Discussion & Conclusion


The original question is if the vaccine is either beneficial, disadvantageous, or neutral in preventing females from contracting covid-19. I can conclude that there IS a difference. I can REJECT the null hypothesis. Since all 3 p-values were below .05 & since the odds ratio contradicts the null hypothesis. Now than, the vaccine POSITIVELY effects females. When females take the vaccine, their odds of contracting covid-19 DECREASES. Thus, I DO recommend that females do take the vaccine since it helps decrease their risks of contracting the deadly disease. 


## LGBTQ+ 

### Null Hypothesis & Alternative Hypothesis, & Methods (Question #3)

$H_0$: There is no correlation between LGBTQ+ members & the vaccine. The vaccine doesn't work on LGBTQ+ members: you have the same likelihood of contracting covid-19 if you do or do not take the vaccine (if you are gay).

$H_A$: There is a correlation between LGBTQ+ members & the vaccine. The drug is either beneficial or disadvantageous in preventing LGBTQ+ members from contracting covid-19. You don't have the same likelihood (if you do or do not take the vaccine) of contracting the disease.

I'll use a bar chart for graphical descriptive results. I'll use a numerical data table for numerical descriptive results. I'll use a chi-squared test & a fisher exact test for inferential results. Than lastly, conclude if I fail to reject or reject the null hypothesis.

### Graphical Descriptive Results 

```{r}
barchartGC(~infected + treatment, data = LGBTQ)
barchartGC(~infected + treatment, data = LGBTQ, type = "percent")
```

The vaccine appears to make LGBTQ+ members more susceptible to covid-19. Since in the first bar chart (in frequency), you can visibly tell that LGBTQ+ members who do take the vaccine contract covid-19 more often. My second bar chart (in percentage), only furthers proves my claim. Since the bar chart value (for infected) increases in the covid-19 department when LGBTQ+ members do take the vaccine by about 70% (if the placebo infected rate is at 15% & if the vaccine infected rate is at 85%). LGBTQ+ members so far, have a disadvantageous result when they take the vaccine. 


### Numerical Descriptive Results

```{r}
table3 <- xtabs(~infected + treatment, data = LGBTQ)
rowPerc(table3)
colPerc(table3)
```

My (row percentages) data table shows that my initial estimate for the increased likelihood of LGBTQ+ members to contract covid-19 with the vaccine being 70% is valid. Since the data table shows that 86.67% of the infected population is vaccinated. Thus, when I subtract the two percentages, the vaccine percentage is higher by 73.34%. Quite close to my estimate of 70%. My (column percentages) data table shows that 3.43% of LGBTQ+ members (who are vaccinated) have contracted covid-19. Thus, 2.91% more of the vaccinated LGBTQ+ member population is infected with covid-19 compared to those who aren't vaccinated. Since when I subtract the two percentages, the vaccine percentage is higher by 2.91%. LGBTQ+ members, so far, continuously show to have a disadvantageous result when they take the vaccine. 

### Inferential Descriptive Results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

The null hypothesis has to be rejected since the chi-squared statistic states that the p-value is 0.0000000132. This number is way below than the .05 mark. The chi-squared testGC also rejects the null hypothesis. It has a chi-squared statistic of 32.30 with a degrees of freedom of 1. I reject the null hypothesis since it states that the p-value is 0. Also way below .05. The fisher exact test states that the p-value is 0.000000001642. The fisher exact test also states that the odds ratio is 0.15 to 1. My alternative hypothesis states that the odds ratio is not equal to 1. Since all 3 p-values are below .05 & the odds ratio contradicts the null hypothesis. I can safely reject the null hypothesis. There is a difference, it being a disadvantageous one for LGBTQ+ members. 

### LGBTQ+ Discussion & Conclusion

The original question is if the vaccine is either beneficial, disadvantageous, or neutral in preventing LGBTQ+ members from contracting covid-19. I can conclude that there IS a difference. I can REJECT the null hypothesis. Since all 3 p-values were below .05 & since the odds ratio contradicts the null hypothesis. Now than, the vaccine NEGATIVELY effects LGBTQ+ members. When LGBTQ+ members take the vaccine, their odds of contracting covid-19 INCREASES. Thus, I DON'T recommend that LGBTQ+ members take the vaccine since it doesn't help decrease their odds of contracting the deadly disease but instead, increase the odds of contracting it. 


## Drug Users



